\documentclass[aps,pre,preprint,groupedaddress]{revtex4}
%\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{epsf,epsfig,graphics}
\usepackage{float}
\usepackage{makeidx}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{amsbsy}
\usepackage{subfigure}
%\usepackage[table,x11names]{xcolor}
%\usepackage{kpfonts}
%\usepackage{xeCJK}
\usepackage{CJKutf8}
\usepackage{color}
\usepackage{morefloats}
\newcommand*{\vv}[1]{\vec{\mkern0mu#1}}
\newcommand{\figwsL}{16cm}
\newcommand{\figws}{7.5cm}
\newcommand{\figwss}{5.cm}
\newcommand{\figw}{10cm}
\newcommand{\figww}{15cm}
%\usepackage{mathaccents}
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

\begin{document}

\title{Linear Response Theory in Noisy Network Dynamics for Nonequilibrium Steady States}
\author{Rong-Chih Chang$^1$\begin{CJK*}{UTF8}{bsmi}(張容誌)\end{CJK*}, Pik-Yin Lai$^{1,2}$%\footnote{pylai@phy.ncu.edu.tw}
\begin{CJK*}{UTF8}{bsmi}(黎璧賢)\end{CJK*}}
\affiliation{$^1$Dept. of Physics and Center for Complex Systems, National Central University, Chung-Li District, Taoyuan City 320, Taiwan, R.O.C.
}
\affiliation{$^2$Physics Division, National Center for Theoretical Sciences, Taipei 10617, Taiwan.}

\begin{abstract}
    Linear response theory predicts how a system responds to small perturbations using information contained in its unperturbed state. Here we develop a computable linear-response framework for noisy network dynamics near stable fixed points. The system consists of coupled stochastic variables with intrinsic node dynamics, weighted possibly directed interactions, and Gaussian white noise. Linearizing the dynamics around the noise-free fixed point yields a Jacobian matrix and a steady-state covariance matrix that satisfy a Lyapunov equation. We derive explicit response formulas for an observable under perturbations of node parameters, connection weights, and noise strengths. The resulting response functions are expressed in terms of correlation functions evaluated in the unperturbed steady state, and we take the equal-time covariance response $\langle \delta x_i \delta x_j \rangle$ for example as a central result. The formulation applies to equilibrium steady states as well as nonequilibrium steady states in which detailed balance is generically broken by asymmetric couplings or heterogeneous noise strengths. We further extend the framework to time-dependent perturbation protocols, yielding transient response formulas for means, forces, and covariance dynamics. Langevin simulations on undirected and directed networks confirm the predicted covariance responses in the linear-response regime.
\end{abstract}

\maketitle

\input{Section/Introduction.tex}
\input{Section/Network_model.tex}
\input{Section/Steady_response.tex}
\input{Section/Response_formula.tex}
\input{Section/Result.tex}
\input{Section/Time-dependent.tex}
\input{Section/Conclusion.tex}
\appendix
\input{Section/Appendix1.tex}
\input{Section/Appendix2.tex}
\input{Section/Appendix3.tex}

\bibliographystyle{apsrev}
\bibliography{ref}

\end{document}
