\section{Spectral solution of the Lyapunov equation and transient protocol integrals}
\quad This appendix collects two technical steps used in the main text: the
spectral solution of the Lyapunov equation and the mode expansion of the
time-dependent convolution integrals in Sec. \ref{sec:time-dependent}.

\subsection*{Spectral solution of the Lyapunov equation}
Consider the Lyapunov equation for the unknown ${\bf K}$ \cite{Behr2019SylvesterLyapunov}
\begin{equation}
{\bf Q}{\bf K}+{\bf K}{\bf Q}^\intercal={\bf S},
\label{lyap}
\end{equation}
where ${\bf Q}$ and ${\bf S}$ are given matrices and ${\bf S}$ is symmetric.
Assume that ${\bf Q}$ is diagonalizable:
\begin{equation}
{\bf Q}={\bf P}\boldsymbol{\Lambda}{\bf P}^{-1},\quad
{\bf P}=({\vec u_1},{\vec u_2},\cdots,{\vec u_N}),\quad
\boldsymbol{\Lambda}=\hbox{diag}(\lambda_1,\lambda_2,\cdots,\lambda_N),
\end{equation}
with ${\bf Q}{\vec u_i}=\lambda_i{\vec u_i}$. Let the columns of
${\bf P}^{-\intercal}$ be the left eigenvectors,
\begin{equation}
{\bf P}^{-\intercal}=({\vec v_1},{\vec v_2},\cdots,{\vec v_N}),
\end{equation}
so that
\begin{equation}
{\vec v_i}^\intercal{\vec u_j}=\delta_{ij},\qquad
\sum_{j=1}^N {\vec u_j}{\vec v_j}^\intercal={\bf I}.
\end{equation}
Defining ${\tilde{\bf K}}={\bf P}^{-1}{\bf K}{\bf P}^{-\intercal}$,
Eq. (\ref{lyap}) becomes
\begin{eqnarray}
\boldsymbol{\Lambda}{\tilde{\bf K}}+{\tilde{\bf K}}\boldsymbol{\Lambda}
&=&{\bf P}^{-1}{\bf S}{\bf P}^{-\intercal},\\
{\tilde{\bf K}}_{ij}
&=&\frac{({\bf P}^{-1}{\bf S}{\bf P}^{-\intercal})_{ij}}{\lambda_i+\lambda_j}.
\label{Ktil}
\end{eqnarray}
Hence
\begin{equation}
{\bf K}={\bf P}{\tilde{\bf K}}{\bf P}^\intercal.
\end{equation}
For the stable fixed points considered here, $\mathrm{Re}\,\lambda_i<0$, so
$\lambda_i+\lambda_j\neq 0$ in Eq. (\ref{Ktil}). Using the biorthogonal
decomposition
\begin{equation}
{\bf Q}=\sum_{m=1}^N \lambda_m {\vec u_m}{\vec v_m}^\intercal,
\end{equation}
the same solution can be written more compactly as
\begin{equation}
{\bf K}=\sum_{m,\ell=1}^N
\frac{({\vec v_m}^\intercal{\bf S}{\vec v_\ell})\,{\vec u_m}{\vec u_\ell}^\intercal}
{\lambda_m+\lambda_\ell}.
\end{equation}

\subsection*{Mode expansion for the time-dependent protocol integrals}
The transient formulas in Sec. \ref{sec:time-dependent} follow from the same
biorthogonal decomposition. Since
\begin{equation}
e^{t{\bf Q}}=\sum_{k=1}^N e^{\lambda_k t}{\vec u_k}{\vec v_k}^\intercal,
\qquad
{\bf Q}^{-1}=\sum_{k=1}^N \frac{{\vec u_k}{\vec v_k}^\intercal}{\lambda_k},
\end{equation}
Eq. (\ref{xi(t)drlinear}) becomes
\begin{equation}
\langle x_i(t)\rangle_{r+\delta r_\alpha(t)}=X_i+(a_\alpha-X_\alpha)\bigg[
\sum_{k=1}^N \frac{u_{ik}v_{\alpha k}}{\lambda_k}\int_0^t dt'
e^{(t-t')\lambda_k}\delta{\dot r}_\alpha(t')
-Q^{-1}_{i\alpha}\delta r_\alpha(t)\bigg],
\label{xi(t)drlinearb}
\end{equation}
where $u_{ik}\equiv({\vec u_k})_i$ and $v_{\alpha k}\equiv({\vec v_k})_\alpha$.

Likewise, for the covariance response,
\begin{eqnarray}
&&\int_0^t dt' \left(
e^{(t-t'){\bf Q}}
\frac{\partial {\bf K}_0}{\partial Q_{\alpha\alpha}}
e^{(t-t'){\bf Q}^\intercal}
\right)_{ij}\delta \dot{r}_\alpha(t') \nonumber\\
&=&\sum_{m,\ell=1}^N
u_{im}u_{j\ell}\left(
{\vec v_m}^\intercal
\frac{\partial {\bf K}_0}{\partial Q_{\alpha\alpha}}
{\vec v_\ell}
\right)I_{m\ell}(t),
\end{eqnarray}
with
\begin{equation}
I_{m\ell}(t)\equiv
\int_0^t dt' \, e^{(t-t')(\lambda_m+\lambda_\ell)}\delta \dot r_\alpha(t').
\end{equation}
For the cosine protocol in Eq. (\ref{dra(t)cos}), this is the same quantity
introduced in Eq. (\ref{Iml}); its explicit piecewise form is given there. The
edge-perturbation formulas in Sec. \ref{sec:time-dependent} follow from the
same expansion, with only the prefactor and the covariance derivative replaced
according to the perturbation under consideration.
