\section{Introduction}
\quad In recent decades, advances in experimental techniques and data acquisition have enabled the recording of high-dimensional time-series data in a wide range of biological, physical, and engineered systems. Many interacting systems in these areas can be described as noisy network dynamics \cite{10.1093/oso/9780198805090.001.0001, doi:10.1137/S003614450342480}, in which heterogeneous components interact through weighted and possibly directed connections while being subject to stochastic fluctuations, where both fluctuation statistics and perturbation responses are experimentally accessible. From a physical perspective, many living systems are maintained through continuous exchanges of energy, matter, and information with their surroundings, and therefore typically operate far from thermodynamic equilibrium \cite{10.1143/PTPS.130.29, sekimoto2010stochastic, Shiraishi2023}. Examples include metabolic processes and molecular motors. In such nonequilibrium systems, energy is continuously consumed and entropy is produced, while detailed balance can be broken by asymmetric interactions or by heterogeneous noise sources acting on different components.\par

One important inverse problem in complex networks is network reconstruction from time-series data. In many situations, the detailed connectivity is not directly accessible, whereas the node dynamics can be measured. Previous studies have shown that fluctuations can encode nontrivial structural information, and that network connectivity can in some cases be inferred from noisy dynamics \cite{PhysRevLett.104.058701, ching2013, ching2015, ching2017, Lai2017, ChengLai2022}. These results highlight the close relation between fluctuations and interactions in networked systems. However, reconstruction alone does not answer a closely related and equally important question: once the parameters of a noisy network are changed, how can one predict the resulting response of the system from properties of its original steady state? In many situations, the detailed microscopic dynamics may be complicated, but the system remains close to a stable fixed point and its fluctuation statistics can still be measured. This motivates a response theory formulated directly for noisy network dynamics near stable fixed points, where one can ask how perturbations of node parameters, couplings, or noise strengths modify steady-state observables.\par

In statistical mechanics, linear response theory (LRT) provides a general framework for relating the response of a system to small perturbations to correlation functions measured in the unperturbed state \cite{RKubo_1966, softmatter, 10.1063/5.0003135}. In equilibrium systems, this connection is expressed by the fluctuation-dissipation theorem (FDT). However, many realistic networked systems do not operate near thermodynamic equilibrium, but rather in nonequilibrium steady states (NESS). Importantly, previous studies have shown that the notion of linear response remains meaningful beyond equilibrium \cite{Seifert_2010, Mehl_2010}. In particular, generalized fluctuation-dissipation relations for NESS have been developed by introducing the correction and experimentally verified. These results suggest that the breakdown of detailed balance does not necessarily invalidate predictive linear response theory.\par

In coupled dynamical systems, local perturbations generally do not remain local. Instead, perturbations of node parameters or interaction strengths can modify fluctuations and correlations throughout the system. In practical terms, this means that local changes, such as node removal, link failure, or targeted interventions, can induce nontrivial changes in global observables \cite{NetworkDisruption}. A predictive framework for such responses is therefore useful not only for understanding noisy interacting systems, but also for assessing the effect of controlled perturbations and disruptions.\par

A general framework that unifies linear response theory in NESS with noisy network dynamics remains relatively underexplored. Although nonequilibrium response has been verified in several low-dimensional, experimentally well-controlled systems, there is still a need for a practical framework that can be computed directly from the unperturbed network and then tested against perturbation experiments or simulations. This need is especially acute for fluctuation observables, since local parameter changes can redistribute correlations across an entire network. The limitation is even more evident when one wishes to describe not only the difference between two steady states, but also the transient relaxation induced by a time-dependent perturbation protocol.\par

In this work, we address this problem for noisy network dynamics near stable fixed points. Each node is driven by Gaussian white noise and interacts with other nodes through weighted directed couplings, while also possessing its own intrinsic dynamics. Linearizing the dynamics around the noise-free stationary state leads to a Jacobian matrix $\mathbf{Q}$ \cite{doi:10.1073/pnas.0506347102} and a steady-state equal-time covariance matrix $\mathbf{K}_0 \equiv \langle \delta \vec{x}\,\delta \vec{x}^{\top} \rangle$ satisfying a Lyapunov equation. Within this structure, we develop a unified response framework for both steady-state and time-dependent perturbations, covering the responses of the mean state, deterministic force, and fluctuation covariance. Our central focus is the response of $\mathbf{K}_0$, for which we derive explicit and computable formulas under perturbations of intrinsic node parameters, connection weights, and noise strengths. The framework applies to both equilibrium steady states and nonequilibrium steady states, where detailed balance is broken by asymmetric couplings or heterogeneous noise amplitudes. In this sense, the contribution of this work is not a general reformulation of NESS response theory, but an operational response framework for linearized noisy networks in which covariance response is the central result, while the responses of means and forces show that the same formalism also captures transient network relaxation more broadly.\par

The paper is organized as follows. In Sec. II, we introduce the noisy network model and its linearization near a stable fixed point. In Sec. III, we establish a linear response framework for both equilibrium and NESS in the linearized Langevin dynamics. In Sec. IV, we derive the general steady-state linear response theory for noisy networks. In Sec. V, we compare the theoretical predictions with Langevin simulations for both equilibrium and NESS cases. In Sec. VI, we extend the formalism to time-dependent perturbations and discuss the corresponding dynamical response.
