\section{Noisy network dynamics and linearization near a stable fixed point}
\quad We consider a network of $N$ nodes, where the state of node $i$ is described by a continuous variable $x_i(t)$. The intrinsic dynamics of each node is governed by a function $f_i(x_i;r_i)$, where $r_i$ denotes a node-specific parameter. Nodes $i$ and $j$ are connected by directed, weighted links described by the matrix ${\bf W}$. The noisy network dynamics are governed by
\begin{equation}
    \dot{x}_i = f_i(x_i;r_i) + \sum_{k\neq i} W_{ik} h(x_i, x_k) + \eta_i(t),\;i=1,2,...,N,
    \label{noisy_network}
\end{equation}
where $\eta_i(t)$ is a zero-mean Gaussian white noise acting on node $i$. The weight matrix can be written as $W_{ij}=g_{ij}A_{ij}$, where $A_{ij}=0$ or 1 is the adjacency matrix and $g_{ij}$ is the connection weight. Both $\mathbf{A}$ and $\mathbf{W}$ may be asymmetric, and self-loops are excluded. We adopt the convention that $A_{ij}$ denotes an edge from node $j$ to node $i$. The coupling function $h(x_i,x_j)$ describes the interaction between nodes $i$ and $j$. Eq. (\ref{noisy_network}) has the mathematical form of overdamped Langevin dynamics on a network. It is convenient to combine the intrinsic and coupling terms into a deterministic force $\vec{F}(\vec{x};\vec{p})$, where $\vec{p}$ denotes a set of control parameters, including the node parameters $r_i$ and the network connection weights $W_{ij}$. The dynamics can then be written as
\begin{equation}
    \dot{\vec{x}} = \vec{F}(\vec{x};\vec{p}) + \vec{\eta}(t),
    \label{deterministic_force}
\end{equation}
with
\begin{equation}
    F_j(\vec{x}; r_j, {\bf W}) = f_j(x_j;r_j) + \sum_{k\neq j} W_{jk} h(x_j,x_k).
\end{equation}
Here the parameters are assumed to be externally tunable and time-independent. This formulation applies to a wide range of noisy network systems. For example, in neuroscience, $x_i(t)$ may represent an average firing rate and $W_{ij}$ the strength and direction of synaptic connections; in ecological systems, $x_i(t)$ may represent the abundance of species $i$, coupled to other species through interaction terms.\par
We assume that the system fluctuates around a noise-free fixed point $\vec{X}$, defined by $\vec{F}(\vec{X}) = 0$. To analyze fluctuations near a stable fixed point, we introduce the deviation
\begin{equation}
    \Delta \vec{x} \equiv \vec{x} - \vec{X},
\end{equation}
and linearize the deterministic force around $\vec{X}$. The Jacobian matrix
\begin{equation}
    \mathbf{Q} = \nabla \vec{F}\big|_{\vec{X}}
\end{equation}
encodes both the network topology and the local dynamics, and is asymmetric in general. Up to linear order in $\Delta \vec{x}$, the stochastic dynamics becomes
\begin{equation}
    \Delta \dot{\vec{x}} = {\bf Q}\Delta \vec{x} + \vec{\eta}(t),
    \label{linearized_dynamics}
\end{equation}
where the Jacobian matrix is given by
\begin{equation}
    Q_{ij} = W_{ij} \partial_2 h(X_i,X_j) + \left[ f_i'(X_i;r_i) + \sum_{m\neq i} W_{im} \partial_1 h(X_i,X_m) \right] \delta_{ij}.
    \label{Qij_definition}
\end{equation}
We assume that the fixed point $\vec{X}$ is linearly stable, such that all eigenvalues of $\mathbf{Q}$ have negative real parts. The noise is taken to be additive Gaussian white noise with covariance matrix
\begin{equation}
    \langle \vec{\eta}(t) \vec{\eta}^\intercal(t') \rangle = \boldsymbol{\sigma} \delta (t-t'),\;\langle \vec{\eta}(t) \rangle = 0,
\end{equation}
where $\boldsymbol{\sigma}$ is chosen to be diagonal. When the system reaches a steady state, the equal-time covariance matrix ${\bf K}_0 \equiv \langle \delta \vec{x} \delta \vec{x}^\intercal \rangle$ satisfies the Lyapunov equation
\begin{equation}
    \boldsymbol{\sigma} = -\mathbf{Q K}_0 - \mathbf{K}_0 \mathbf{Q}^\intercal.
    \label{Lyapunov_eq}
\end{equation}
This relation directly links the steady-state fluctuation statistics to the linearized drift matrix and noise matrix, and can be viewed as a multidimensional version of the FDT \cite{lai2026steadystate}. Given $\mathbf{Q}$ and $\boldsymbol{\sigma}$, $\mathbf{K}_0$ can be obtained by solving the Lyapunov equation. The dynamics is also known as a multivariate Ornstein-Uhlenbeck process, and its steady-state probability distribution is Gaussian:
\begin{equation}
    \psi_{ss}(\delta \vec{x}; \vec{p}) = \frac{1}{(2\pi)^{N/2}\sqrt{|{\bf K}_0|}} e^{-\frac{1}{2} \delta \vec{x}^\intercal {\bf K}^{-1}_0 \delta \vec{x}},
    \label{steady-state prob}
\end{equation}
which can be written in terms of the effective potential $\Phi \equiv \frac{1}{2} \delta \vec{x}^\intercal {\bf K}^{-1}_0 \delta \vec{x}$, where $\delta \vec{x} \equiv \vec{x} - \langle \vec{x} \rangle$. For the linearized dynamics, the steady-state mean coincides with the fixed point, $\vec{X}=\langle \vec{x} \rangle$, so that $\delta \vec{x} = \Delta \vec{x}$. The distinction between these two quantities for nonlinear forces is discussed in Appendix~A. The ensemble average of an observable $B(\vec{x}; \vec{p})$ is given by $\langle B \rangle = \int B(\vec{x}; \vec{p}) \psi_{ss}(\vec{x}; \vec{p}) d\vec{x}$.\par
However, the existence of a stationary Gaussian distribution does not by itself imply equilibrium. This distinction can also be detected from the time-lagged covariance $\mathbf{K}_{\tau} \equiv \langle \delta \vec{x}(t+\tau) \delta \vec{x}^\intercal(t) \rangle$. The condition $\mathbf{K}_\tau = \mathbf{K}^{\intercal}_\tau$ for all $\tau$ is the time-reversal symmetry condition associated with detailed balance in linear Gaussian dynamics. Using
\begin{equation}
    \delta \vec{x}(t+\tau) = e^{\tau \mathbf{Q}} \delta \vec{x}(t) + \int_t^{t+\tau} ds\, e^{(t+\tau-s)\mathbf{Q}} \vec{\eta}(s),
\end{equation}
one obtains $\mathbf{K}_\tau = e^{\tau \mathbf{Q}}\mathbf{K}_0$, where $s > t$, so $\vec{\eta}(s)$ does not correlate with $\vec{x}(t)$. The equilibrium condition is therefore
\begin{equation}
    \mathbf{Q}\mathbf{K}_0 = \mathbf{K}_0\mathbf{Q}^\intercal,
\end{equation}
or equivalently
\begin{equation}
    \mathbf{Q}\boldsymbol{\sigma} = \boldsymbol{\sigma}\mathbf{Q}^\intercal.
    \label{equilibrium_condition}
\end{equation}
Thus $\mathbf{Q}\boldsymbol{\sigma}$ must be symmetric. This includes the case of symmetric $\mathbf{Q}$ with a uniform diagonal noise matrix $\boldsymbol{\sigma}\propto \sigma_0 \mathbf{I}$, but it is more general and does not require $\mathbf{Q}$ itself to be symmetric for special choices of $\boldsymbol{\sigma}$. For diagonal $\boldsymbol{\sigma}$, the condition becomes
\begin{equation}
    Q_{ij} \sigma_{jj} = Q_{ji} \sigma_{ii}.
    \label{equilibrium_condition_diagonal}
\end{equation}
Although asymmetric couplings or heterogeneous noise generally break detailed balance, the equilibrium criterion is fully determined by Eq. (\ref{equilibrium_condition}). In particular, there can still be special diagonal choices of $\boldsymbol{\sigma}$ that satisfy Eq. (\ref{equilibrium_condition_diagonal}) even when $\mathbf{Q}$ is asymmetric. Beyond the linearized regime, the full nonlinear dynamics still admits a steady-state form $\psi_{ss}(\vec{x}) \propto e^{-\Phi(\vec{x})}$, but $\Phi$ is generally not quadratic and the local peak of $\psi_{ss}$ need not coincide exactly with either the noise-free fixed point or the steady-state mean \cite{PhysRevE.84.061106}. Near a stable fixed point, however, the leading contribution reduces to the quadratic form in Eq. (\ref{steady-state prob}), so the linearized theory captures the local steady-state structure and response to leading order.\par
From the reconstruction perspective, a fuller derivation of the relations among $\mathbf{K}_\tau$, $\mathbf{Q}$, and $\boldsymbol{\sigma}$ is summarized in Appendix~A.
