\section{Explicit Response Formula For the Covariance Matrix}
\quad We now present the explicit response formulas corresponding to the steady-state linear-response theory developed in the previous section. Throughout this section, the formulas are understood within the local Gaussian approximation associated with the linearized dynamics around a stable fixed point. Equivalently, they give the exact first-order susceptibilities of the linearized Ornstein-Uhlenbeck process and provide the leading fixed-point approximation for the original nonlinear noisy network. As discussed in Appendix~A, $\Delta \vec{x}$ and $\delta \vec{x}$ need not coincide exactly for general nonlinear noisy systems, but they agree to leading order near the fixed point and coincide exactly for linear dynamics. As a representative example, we consider the response of this equal-time covariance matrix with respect to a parameter change in $\vec{p}$, where $\vec{p}$ may denote $\vec{r}$, $\mathbf{W}$, or $\boldsymbol{\sigma}$. The observable is then $B_{ij}(\vec{x}; \vec{p})=\delta x_i \delta x_j$. We also present the corresponding response functions for the different types of parameter perturbations. From the general formula (\ref{response_function2}), the response function with respect to $p_\alpha$ is
\begin{equation}
    \chi_{ij\alpha} = \langle \partial_p (\delta x_i \delta x_j) \rangle_{ss, \vec{p}} + \langle \delta (\delta x_i \delta x_j) \delta \mathscr{F}_\alpha \rangle_{ss, \vec{p}}.
\end{equation}
The observable can be written as $\delta x_i \delta x_j= [x_i - \langle x_i \rangle][x_j - \langle x_j \rangle]$. Then
\begin{equation}
    \left\langle \frac{\partial B_{ij}}{\partial p_\alpha} \right\rangle = -\frac{\partial \langle x_i \rangle}{\partial p_\alpha} \langle \delta x_j \rangle - \frac{\partial \left\langle x_j \right\rangle}{\partial p_\alpha} \langle \delta x_i \rangle.
\end{equation}
Since $\langle \delta x_i \rangle = \langle \delta x_j \rangle = 0$, the first term vanishes. The second term in Eq. (\ref{response_function2}) can be written as
\begin{eqnarray}
    \left\langle \delta (\delta x_i \delta x_j) \delta \mathscr{F}_\alpha \right\rangle_{ss, \vec{p}} &=& \left\langle (\delta x_i \delta x_j) \mathscr{F}_\alpha \right\rangle_{ss, \vec{p}} - \left\langle \delta x_i \delta x_j \right\rangle_{ss, \vec{p}} \left\langle \mathscr{F}_\alpha \right\rangle_{ss, \vec{p}}\nonumber\\
    &=& \langle (\delta x_i \delta x_j) \mathscr{F}_\alpha \rangle_{ss, \vec{p}} - (\mathbf{K}_0)_{ij} \langle \mathscr{F}_\alpha \rangle_{ss, \vec{p}}.
\end{eqnarray}
The explicit conjugate-force formulas and the resulting covariance-response identities for node, edge, and noise perturbations are derived in Appendix~B; see in particular Eqs. (\ref{chiKodr_ss_appB}), (\ref{chiKodW_ss_appB}), and (\ref{chiKosigma_ss_appB}). Those appendix expressions give the general first-order covariance responses within the linearized Gaussian description. The simplified formulas written below apply only under the stated model assumptions.

For the benchmark simplifications used below and later validated in Sec.~V, we restrict attention to the two model classes introduced there: the linear restoring-force system $f_i(x_i)=-r_i(x_i-1)$ with diffusive coupling $h(x_i,x_j)=x_j-x_i$, and the logistic benchmark $f_i(x_i)=r_i x_i(1-x_i)$ with the same diffusive coupling evaluated at the synchronized fixed point $X_i=1$. Under these benchmark assumptions, one has $h(X_i,X_j)=0$, $\partial_1 h(X_i,X_j)=-1$, $\partial_2 h(X_i,X_j)=1$, and
\begin{equation}
    Q_{\alpha\alpha}=-r_\alpha-\sum_j W_{\alpha j},\qquad Q_{\alpha\beta}=W_{\alpha\beta}\;\;(\alpha\neq\beta).
\end{equation}
For the logistic benchmark at $X_i=1$, one also has $\partial f_\alpha/\partial r_\alpha|_{\vec X}=0$ and $\partial f'_\alpha/\partial r_\alpha|_{\vec X}=-1$.
\subsection{Response due to Change in the Intrinsic Node Parameter $r$}
\quad We first consider a perturbation of a single node parameter $r_\alpha$, where $\alpha$ denotes the index of the perturbed node. The steady-state response of the equal-time covariance matrix is
\begin{equation}
    \chi_{ij\alpha} = \langle \delta x_i \delta x_j \mathscr{F}_\alpha \rangle - (\mathbf{K}_0)_{ij} \langle \mathscr{F}_\alpha \rangle.
    \label{response_Ko_r}
\end{equation}
Under the benchmark assumptions stated above, the perturbation enters the Jacobian through the diagonal element $Q_{\alpha\alpha}$, so Eq. (\ref{response_Ko_r}) reduces to
\begin{equation}
    \chi_{ij\alpha}^{(ss)} = -\left(\frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\alpha}} \right)_{ij}.
    \label{chiKodr_ss}
\end{equation}
In these cases, the covariance response is directly controlled by the local restoring parameter and the total incoming coupling associated with node $\alpha$. The derivative of $\mathbf{K}_0$ can be obtained by differentiating both sides of the Lyapunov equation (\ref{Lyapunov_eq}), as discussed later in this section.
\subsection{Response due to Change in the Connection Weight $W$}
\quad We next consider a perturbation of a directed connection weight $W_{\alpha\beta}$. The corresponding response function of the covariance matrix is
\begin{equation}
    \chi_{ij(\alpha\beta)} = \langle \delta x_i \delta x_j \mathscr{F}_{\alpha\beta} \rangle - (\mathbf{K}_0)_{ij} \langle \mathscr{F}_{\alpha\beta} \rangle.
    \label{response_Ko_W}
\end{equation}
Under the same benchmark assumptions, the perturbed coupling does not shift the fixed point at leading order because $h(X_\alpha, X_\beta)=0$. The terms proportional to $h(X_\alpha,X_\beta)$ in Appendix~B then vanish, and Eq. (\ref{response_Ko_W}) reduces to
\begin{equation}
    \chi_{ij(\alpha\beta)} = \partial_1 h(X_\alpha,X_\beta) \left( \frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\alpha}} \right)_{ij} + \partial_2 h(X_\alpha,X_\beta) \left( \frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\beta}} \right)_{ij},
    \label{chiKodW_ss}
\end{equation}
which makes the role of the perturbed connection especially transparent. In this restricted setting, the response is determined by two contributions: the change in the local diagonal term $Q_{\alpha\alpha}$ and the change in the directed off-diagonal coupling $Q_{\alpha\beta}$. For an undirected network, since $W_{\alpha\beta}=W_{\beta\alpha}$, the total response is the sum of the two directed contributions associated with $W_{\alpha\beta}$ and $W_{\beta\alpha}$.
\subsection{Response due to Change in the Noise Strength $\sigma$}
\quad Finally, we consider a perturbation of a noise strength. In contrast to the previous two cases, a change in $\sigma_{\alpha\alpha}$ does not alter the deterministic force and therefore does not shift the fixed point. The corresponding response function is
\begin{equation}
    \chi_{ij(\alpha\alpha)} = \langle \delta x_i \delta x_j \mathscr{F}_{\alpha\alpha} \rangle - (\mathbf{K}_0)_{ij} \langle \mathscr{F}_{\alpha\alpha} \rangle = \left( \frac{\partial \mathbf{K}_0}{\partial \sigma_{\alpha\alpha}} \right)_{ij}.
\end{equation}
This result emphasizes that a perturbation of the noise strength affects the steady-state fluctuations directly, without changing the deterministic fixed point.

\subsection{Evaluation of $\partial \mathbf{K}_0/\partial Q_{\alpha\beta}$ and $\partial \mathbf{K}_0/\partial \sigma_{\alpha\alpha}$}
\quad Within the same linearized Ornstein-Uhlenbeck description, the derivative of $\mathbf{K}_0$ with respect to $\mathbf{Q}$ or $\boldsymbol{\sigma}$ is determined from the steady-state Lyapunov equation because $\mathbf{K}_0$ depends on both $\mathbf{Q}$ and $\boldsymbol{\sigma}$. Appendix~A summarizes the related relations among $\mathbf{K}_\tau$, $\mathbf{Q}$, and $\boldsymbol{\sigma}$, while the derivatives used here follow directly from differentiating the steady-state Lyapunov equation. The derivatives of $\mathbf{K}_0$ with respect to the matrix elements of $\mathbf{Q}$ can be obtained by differentiating the Lyapunov equation:
\begin{equation}
    \mathbf{Q} \frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\beta}} + \frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\beta}} \mathbf{Q}^\intercal = \mathbf{M}^{(\alpha,\beta)},\; (\mathbf{M}^{(\alpha,\beta)})_{ij} \equiv -(\mathbf{K}_0)_{i\beta} \delta_{j\alpha} - (\mathbf{K}_0)_{j\beta} \delta_{i\alpha}.
    \label{derivative_Ko}
\end{equation}
Likewise, the derivative of the inverse covariance matrix gives
\begin{equation}
    \mathbf{K}_0 \frac{\partial \mathbf{K}^{-1}_0}{\partial Q_{\alpha\beta}} \mathbf{K}_0 = - \frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\beta}}\;\text{or}\;\mathbf{K}^{-1}_0 \frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\beta}} \mathbf{K}^{-1}_0 = - \frac{\partial \mathbf{K}^{-1}_0}{\partial Q_{\alpha\beta}}.
    \label{inverse_partialKo}
\end{equation}
Higher-order derivatives can be obtained in a similar manner. For the derivative with respect to $\sigma_{\alpha\alpha}$, we have
\begin{equation}
    \mathbf{Q}\frac{\partial\mathbf{K}_0}{\partial\sigma_{\alpha\alpha}} + \frac{\partial\mathbf{K}_0}{\partial\sigma_{\alpha\alpha}}\mathbf{Q}^\intercal = -\mathbf{E}^{(\alpha\alpha)},\;\text{where}\;E^{(\alpha\alpha)}_{ij}=1\;\text{for}\;i=j=\alpha.
    \label{partialKo_sigma}
\end{equation}
Moreover, the above formulas simplify further when the network satisfies the equilibrium condition $\mathbf{Q}\boldsymbol{\sigma}=\boldsymbol{\sigma}\mathbf{Q}^\intercal$. In that case, the covariance matrix satisfies $\mathbf{K}_0 = -\frac{1}{2} \mathbf{Q}^{-1} \boldsymbol{\sigma}$, and therefore
\begin{equation}
    \left( \frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\beta}} \right)_{ij} = -Q^{-1}_{i\alpha} (\mathbf{K}_0)_{j\beta} = \frac{1}{2} Q^{-1}_{i\alpha} (\mathbf{Q}^{-1}\boldsymbol{\sigma})_{j\beta}.
\end{equation}
If the nodes share the same noise strength $\boldsymbol{\sigma}=\sigma \mathbf{I}$, the formula is further simplified
\begin{equation}
    \left( \frac{\partial \mathbf{K}_0}{\partial Q_{\alpha\beta}} \right)_{ij} = \frac{\sigma}{2} Q^{-1}_{i\alpha} Q^{-1}_{j\beta}.
\end{equation}
This equilibrium derivative identity is especially useful because the dependence of $\mathbf{K}_0$ on $\mathbf{Q}$ is explicit. However, the parameter-response formulas obtained from it still depend on how a given perturbation modifies $\mathbf{Q}$ or $\boldsymbol{\sigma}$. Under the equilibrium version of the benchmark assumptions above, a node perturbation changes only $Q_{\alpha\alpha}$ according to $\delta Q_{\alpha\alpha}=-\delta r_\alpha$, and an edge perturbation changes $Q_{\alpha\alpha}$ and $Q_{\alpha\beta}$ according to $\delta Q_{\alpha\alpha}=-\delta W_{\alpha\beta}$ and $\delta Q_{\alpha\beta}=\delta W_{\alpha\beta}$. The response function of $\mathbf{K}_0$ to a node-parameter perturbation then simplifies to
\begin{equation}
    \chi_{ij(\alpha)} = -\frac{\sigma}{2} Q^{-1}_{i\alpha} Q^{-1}_{j\alpha},
    \label{chiKodr_eq}
\end{equation}
and the response function for a change in a connection weight becomes
\begin{equation}
    \chi_{ij(\alpha\beta)} = -\frac{\sigma}{2} Q^{-1}_{i\alpha} Q^{-1}_{j\alpha}+\frac{\sigma}{2} Q^{-1}_{i\alpha} Q^{-1}_{j\beta},
    \label{chiKodW_eq}
\end{equation}
However, the simplified equilibrium expression applies only to perturbation that preserves detailed balance. Note that a noise perturbation usually breaks detailed balance. For a system that is initially at equilibrium, the perturbed state therefore need not remain in equilibrium. The response formula for noise perturbations should then be derived from Eq. (\ref{partialKo_sigma}). These results provide a simple reference form for comparison with the nonequilibrium steady-state response.
